Supporting Information for ``The dynamics of starvation and recovery''

### 2017-06-09 Manuscript/Nature-Submission/Revised-Supplement-3.tex
Nature submission

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability. Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources. The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.

In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citenum{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment.

Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification. As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$. Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
\end{split}
\end{align}
%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.

>>>>

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citenum{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)

%\begin{align}
%\begin{split}
%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
%\end{split}
%\end{align}
%
%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability. Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources. The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
%
%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification. As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$. Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to

%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
